```{r, echo=FALSE}
term <- function(x) {
  if (knitr::is_latex_output()) {
    sprintf("\textcolor{%s}{%s}", "blue", x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", "blue",
      x)
  } else x
}
```

```{r, echo=FALSE}
preq <- function(x) {
  if (knitr::is_latex_output()) {
    sprintf("\textcolor{%s}{%s}", "magenta", x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", "fuchsia",
      x)
  } else x
}
```

# Fibre Bundles {#fibre-bundles}

## Construction

Let $M, F$ be manifolds and $E$ be a `r term("fibre bundle")` with `r term("base")` $M$ and `r term("fibre")` $F$, or a manifold endowed with a surjective projection $\pi : E \to M$ such that $M$ admits an open covering $\{U_\alpha\}$ for which each $U_\alpha$ has a diffeomorphism $\phi_\alpha : \pi^{-1}(U_\alpha) \to U_\alpha \times F$ acting by $p \to (\pi(p), \xi_\alpha(p))$ for $p \in P$ and some $\xi_\alpha : U_\alpha \to \diff(F)$. This implies $\pi^{-1}(m)$ is diffeomorphic to $F$; we say $E$ is locally a product of $M$ and $F$ and that the $(U_\alpha, \phi_\alpha)$ is a `r term("local trivialization")`. On $U_\alpha \cap U_\beta$ we have functions $\phi_{\alpha\beta} = \phi_\alpha \circ \phi_\beta^{-1} : (U_\alpha \cap U_\beta) \times F \to (U_\alpha \cap U_\beta) \times F$ given by $(m, x) \mapsto (m, \xi_{\alpha\beta}(m)(x))$ for some $\xi_{\alpha\beta}(m) \in \diff(F)$ called `r term("transition functions")`. Sometimes we require $\xi_{\alpha\beta}(m) \in G$, a `r term("topological group")` acting on $F$ on the left by diffeomorphisms (i.e., a subgroup of $\diff(F)$). If a fibre bundle's local trivializations satisfy this maximally, we say $E$ is a `r term("$G$-bundle")`, and that $G$ is the `r term("structure group")`. We note that $\xi_{\alpha\alpha} = 1$, $\xi_{\alpha\beta} = \xi_{\beta\alpha}^{-1}$, and the `r term("cocycle condition")` $\xi_{\alpha\beta} \circ \xi_{\beta\delta} \circ \xi_{\delta\alpha} = 1$ holds on triple overlaps. Fibre bundles are uniquely determined by the base manifold and the transition functions. Given a manifold $N$ and a map $g : N \to M$, the pullback bundle $g^*E$ is the subset of $N \times E$ such that $g \circ \proj{1} = \pi \circ \proj{2}$ with projection $\proj{1} : g^*E \to N$. A `r term("vector bundle")` is a fibre bundle whose fibres are vector spaces and whose local trivializations are fibre-wise linear isomorphisms.

## Lie Groups

A `r term("Lie group")` $G$ is a smooth manifold with group structure such that the binary operation is smooth.[^com:smooth_inverse] 
Elements $g \in G$ induce $R_g, L_g, A_g \in \diff(G)$ by $R_g : h \mapsto hg$, $L_g : h \mapsto gh$, and $A_g = L_g \circ R_{g^{-1}} : h \mapsto ghg^{-1}$. 

## Principal Bundles

Let $G$ be a Lie group and $P$ be a `r term("principle $G$-bundle")` with base $M$, or a $G$-bundle over $M$ with fibre $G$ and transition functions given by left multiplication. Because left and right multiplication commute, we have an invariant right action of $G$ on $P$[^com:pb_invariant_right_action]. Equivalently, a principle $G$-bundle $P$ is a fibre bundle with a `r preq("regular")` smooth right action by a Lie group $G$ that preserves fibres and the $\xi_\alpha$ are $G$-equivariant. It follows that $P/G = M$[^com:pb_def] and $E$ admits a local trivialization with $M \times G$.

## Associated Bundles

## Sections

### Vector Fields

### Tensor Fields

### Differential Forms

[^com:smooth_inverse]: Some authors require the inverse map $g \mapsto g^{-1}$ to be smooth as well, but this follows from the smoothness of the binary operation.

[^com:pb_invariant_right_action]: If $p \in \pi^{-1}(U_\alpha)$ and $\phi_\alpha(p) = (m,h)$, then $pg = \phi_\alpha^{-1}(m,hg)$.

[^com:pb_def]: The second definition admits an exchange between fibre-preservation and $P/G = M$
